Response diversity and food-web structure drive stability and stability-richness relationships in complex communities

### 2024-04-24 paper/manuscript.Rmd
Update manuscript results, discussion and methods

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
#csl: nature.csl
csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

- Putative titles:
  - Community stability is driven by biodiversity and response diversity in
    complex stochastic food-webs
  - Response diversity and food-web structure drive stability and
    stability-richness relationships in fluctuating environments

- Two strong results:
  - Community stability raises primary from richness and average trophic level
    but average interaction strength and connectance have little effects
  - Response diversity and community structure interact strongly to determine
    the sign of the relationship between community stability and species
    richness

# Abstract (150/150 words)

The diversity of species responses to environmental stochasticity is key to
understand temporal community stability and stability-diversity relationships.
While we know quite well how response diversity and species richness drive
community stability in plant communities, whether these effects also operate
in complex food-webs remains an open question. Using a stochastic Bioenergenetic
model, we show that response diversity and species richness are also strong drivers
of overall community stability in complex food-webs. In contrast, we found that
interaction strength and connectance had little effects on temporal stability
overall. Negative stability-richness relationships always emerged in the absence
of response diversity, but we found that interaction strength, connectance,
average trophic level strongly interact with response diversity to determine the
sign of stability-diversity relationships. Our study disentangles the complex
effects of food-web structure on community stability and its pathway, showing
that response diversity is a mechanism that generate positive
stability-diversity relationships, beyond plant communities.

@@

Predicting how ecological communities respond ever-growing anthropogenic
pressures is a central focus in Ecology [@newbold_global_2015;
@services_summary_2019]. Most approaches so-far involve static environment and
simplified species assemblages, overlooking environmental stochasticity and the
complex structure of ecological communities. However, the effects of
stochasticity is much more than noise and lead to emerging properties at the
community level. The key mechanism is the way that species respond to
environmental stochasticity, the so called response diversity
[@elmqvist_response_2003; @loreau_species_2008; @ives_stability_2000;
@ross_how_2023]. Response diversity, even considered as pure random species
fluctuations, generate asynchrony in species fluctuations at the community
level. Coined as "statistical averaging", "portfolio effects" or "insurance
hypothesis", it is responsible for the positive relationships between temporal
stability of community and species richness in plant communities
[@doak_statistical_1998; @yachi_biodiversity_1999; @tilman_biodiversity_2006;
@zhao_biodiversity_2022]. But species responses to environmental fluctuations is
also constrained by species interactions, and so community structure
[@loreau_species_2008; @vasseur_environmental_2007; @thebault_trophic_2005;
@thebault_relationship_2006]. A pioneer example by @vasseur_environmental_2007,
in which they present a counter-intuitive example, where the lack of response
diversity can dampen species oscillations and then increase the temporal
stability in a four species diamond food-web [@mccann_weak_1998]. Despite its
overarching importance on community stability and stability-richness
relationships, the effects of stochasticity and response diversity in complex
communities, such as food-webs, remain mostly unknown.

Understanding emerging temporal community stability (hereafter referred to as
community stability) requires scaling down the focus to population scale
[@doak_statistical_1998; @thibaut_understanding_2013; @zhao_biodiversity_2022].
Community stability, measured as the inverse of the coefficient of variation,
can be partitioned into average population stability and asynchrony
[@thibaut_understanding_2013; @zhao_biodiversity_2022], which respectively
measures the amplitude of species fluctuations relative to their
biomass/abundance and how much species fluctuations’ are compensating each other
at the community level. Asynchrony can then stem from statistical averaging of
independent species fluctuations ["portfolio effects", @doak_statistical_1998;
@tilman_ecological_1999] or compensatory dynamics resulting from response
diversity to environmental fluctuations and species interactions
[@zhao_biodiversity_2022]. The body of theory of BEF, focusing mainly on plants,
showed that response diversity and species richness give rise to asynchrony and
thereby increase community stability, despite that species richness can decrease
population stability in the same time [@tilman_biodiversity_2006;
@houlahan_negative_2018]. On the other side, food-web theory focused more on
population stability. It shows that the strength [@may_will_1972;
@mccann_weak_1998; @thebault_trophic_2005] and distribution [@may_will_1972;
@angelis_stability_1975; @mccann_weak_1998; @mccann_reevaluating_1997] of
species interactions determine population stability ("complexity-stability"
theory) and stability-richness relationship [two trophic level systems,
@thebault_trophic_2005; @ives_stability_2007], while the number of trophic
levels determine population stability [@shanafelt_stability_2018]. Food-web
structure also interacts with response diversity to determine population
stability in diamond shaped food-webs [@ripa_food_2003;
@vasseur_environmental_2007; @ruokolainen_environmental_2013]. So far, food-web
theory overlooked the effects of food-web structure on response diversity and
asynchrony while BEF theory overlooked the complex effects of food-web structure
on population stability and asynchrony. Two keys advances allow to overcome that
limitation: the development of the Bioenergenetic model that enable to model
arbitrary complex communities [@yodzis_body_1992; @brose_allometric_2006] and
the development of statistical tools that enables to partition community
stability into population stability and asynchrony [@thibaut_understanding_2013;
@zhao_biodiversity_2022].

Here, we investigated the effects of species richness and community structure on
community stability and on stability-richness relationship in complex stochastic
food-web. To do so, we developed an extension of the Bioenergetic model
[@brose_allometric_2006; @lajaaiti_ecologicalnetworksdynamicsjl_2024] which
includes environmental stochasticity and response diversity
[@vasseur_environmental_2007]. We modelled food-web varying in size (species
richness XX) and structure (Connectance XX, Predator-Prey Mass Ratio: XX) to
create a diversity of community structure. We additionally varied the
variability of environmental stochasticity (standard deviation of mortality
rates) and response diversity ($1-\rho$, $\rho$ being the correlation among
species mortality rates). We performed XXXXX simulations over a minimum 2000
timesteps (Table S1) and collected the 500 last timesteps to compute community
stability and structure. Using a structural equation model, we deciphered the
ways in which stressors (environmental stochasticity, decreasing response
diversity) and community structure (species richness, average interaction
strength, average trophic level, connectance) determine community stability. In
a second time, we used a linear model to investigate how community structure
interact with response diversity to determine the sign of the stability-richness
relationship.

---
#Here we deliver inference from modelling of complex biodiverse communities on
#how food-web structure (connectance, average interaction strength, average
#trophic level), response diversity, and environmental stochasticity combine to
#shape temporal stability of community biomass, as well as its partition into
#asynchrony and population stability. We then assess how food-web complexity and
#response diversity shape stability-richness relationships.

#Environmental forcing (such as food availability and
#warming) and land-use intensity was reported to synchronise population
#fluctuations [@benton_population_2001; @bluthgen_land_2016; @song_reduced_2015;
#@liu_aridity_2023] while habitat degradation tends to decrease population
#stability [@olivier_urbanization_2020]. Understanding the effects of global
#change on community temporal stability therefore require to account for their
#simultaneous effects on community and population scales [@kefi_advancing_2019;
#@simmons_refocusing_2021].

#Increasing species richness has been reported to decrease population stability
#in experimental grasslands [@tilman_biodiversity_2006;
#@roscher_identifying_2011], while increasing population stability in natural
#settings [@houlahan_negative_2018] [add mechanisms?]. In turn, higher species
#richness often results in higher asynchrony, through portfolio effects (i.e.
#statistical averaging effects due to independent species fluctuations) and
#compensatory dynamics (due to species interactions and response diversity of
#species due to the environment) [@loreau_species_2008;
#@thibaut_understanding_2013]. Systematic increase in asynchrony with species
#richness is the mechanism that drives reported positive relationships between
#community stability and species richness [ref]. Anthropogenic pressures, such as
#eutrophication, has been reported to weakens the relationship between species
#richness and asynchrony in grasslands [@hautier_eutrophication_2014] but such
#empirical evidence remains scarce. However, stressors such as biotic
#homogenization [@buisson_toward_2013; @muthukrishnan_invasive_2020;
#@saladin_rapid_2020] and environmental forcing that decrease species response
#diversity should dampens the relationships between asynchrony and species
#richness [@ives_stability_2000; @thibaut_understanding_2013] due to a decrease
#in portfolio effects and compensatory effects. [Talk about the effects of
#evenness in temporal variability and evenness in abundance? e.g.,
#@zhao_biodiversity_2022; @thibaut_understanding_2013]. Because stressors can
#affect both community stability per se and how community stability scales with
#species richness, improving our understanding require to study both
#[@ives_stability_2007; @hautier_anthropogenic_2015].

#including levels of generalism, the distribution of interaction strengths and
#the prevalence of omnivory, can also influence stability. The degree of
#generalism of predator was shown to modulate stability-richness relationships
#[@thebault_trophic_2005], and generalist predators can synchronise the dynamics
#of their preys[@raimondo_interspecific_2004]. The strength and distribution of
#interaction strength was also shown to be strong driver of community stability,
#such as stronger interaction strengths can be destabilizing [@may_will_1972;
#@mccann_weak_1998]. Omnivory was also shown to drive higher population stability
#[@mccann_reevaluating_1997]. The effects of food-web structure on ecosystem
#stability has been mostly assessed with small modules [but see
#@eschenbrenner_diversity_nodate], and we know very little how food-web structure
#affects the relationships between species richness and temporal stability in
#complex food-webs. So understanding the effects of stressors on community
#stability and on community stability-richness relationships necessitate to
#account for food-web structure.
---

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017]. This model
is widely used to simulate the dynamics of complex ecological communities
because of the simplicity  and because it can be parametrized using  the
metabolic theory of ecology [@brown_toward_2004]. The model describes species
biomass dynamics over time with two master equations (eq. \@ref(eq:dbdt)), one
for the primary producers (eq. \@ref(eq:prod)) and one for the consumers (eq.
\@ref(eq:prod)):

\begin{subequations}
    \label{eq:dbdt}

    \begin{equation}
        \label{eq:cons}
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i - d_i B_i
    \end{equation}

    \begin{equation}
    	\label{eq:prod}
    	\frac{dB_i}{dt} =
	r_i B_i G_i
	- \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - d_i B_i
    \end{equation}
\end{subequations}

@@

\begin{subequations}
    \label{eq:func}
    \begin{equation}
        \label{eq:growth}
        G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
    \end{equation}

    \begin{equation}
    	\label{eq:holling}
        F_{ij} = \frac{\omega_{ij} B_j^h}
        {B_0^h + c_i B_i
        	+ \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
    \end{equation}
\end{subequations}

@@

\begin{equation}
\begin{aligned}
SAE = SAE_{even} \times EVN
(\#eq:sae)
\end{aligned}
\end{equation}

@@

## Drivers of stability

```{r}
tar_load(c(sem_main_text_eff_tab, sem_simplified_eff_tab))

>>>>

---
title: Response diversity and food-web structure drive stability and stability-richness relationships in complex communities
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

# Abstract (179/200 words)

Global environmental change constitutes a suite of major threats to biodiversity
and ecosystem functioning. These threats can materialised via changes to the
stability of communities and the services they provide. Despite independent
advances in understanding the impact on stability of community size and
structure, synchrony and responses to environmental variation, there remains no
core theory connecting these key processes. Developing this theory requires
combining concepts from diversity - stability relationships, synchrony and the
portfolio effect, and the emerging field of response diversity. Using a
stochastic Bioenergenetic food-web model of 20-50 interacting species, we show
that response diversity and species richness are also strong drivers of overall
community stability in complex food-webs. We further show that connectance and
average trophic level consistently dampened asynchrony but that average trophic
level stabilises population stability. Positive stability-richness relationships
always emerged only in presence of response diversity, but interaction strength,
connectance, average trophic level strongly interact with response diversity to
determine the sign of stability-diversity relationships. Our study disentangles
the complex effects of food-web structure on community stability and its
pathway, showing that response diversity is a mechanism that generates positive
stability-diversity relationships, beyond plant communities.

@@

Ecological communities are facing ever-growing environmental variability caused
by climate and environmental change[@lee_ipcc_2023;
@simmons_refocusing_2021]. Decreases in the temporal stability of communities
are increasingly documented[@olivier_urbanization_2020;
@de_boeck_patterns_2018; @hansen_climate_2013; @bluthgen_land_2016; @harris_biological_2018], raising
concerns about the long-term consequences at the ecosystem scale. Repeated calls
have been made to clarify the role of environmental stochasticity in driving the
temporal stability of ecological communities[@yang_predictability_2019]. It
remains challenging because species have a diversity of response to
environmental change [@elmqvist_response_2003; @loreau_species_2008;
@ives_stability_2000; @ross_how_2023] (“response diversity”) and they are
entangled in complex communities such as food-webs, that both ultimately drive
community stability[@may_will_1972; @angelis_stability_1975; @mccann_weak_1998;
@mccann_reevaluating_1997; @thebault_stability_2010] and stability-richness
relationships[@thebault_trophic_2005; @thebault_relationship_2006]. Despite the
overarching effects of response diversity and food-web structure on temporal
community stability, they have been so far studied in isolation.

The temporal stability of ecological communities stems from average population
stability and asynchrony of population fluctuations[@thibaut_understanding_2013;
@zhao_biodiversity_2022]. While higher environmental stochasticity/variability
decrease population stability by increasing the amplitude species oscillations,
higher species response diversity increases asynchrony of species
fluctuations[@doak_statistical_1998; @loreau_species_2008;
@thibaut_understanding_2013]. This is increasingly well-established in
simplified species assemblages, where high response diversity increases
asynchrony and is the determinant of positive stability-diversity relationships
across many taxa[@tilman_biodiversity_2006; @olivier_urbanization_2020;
@zhao_biodiversity_2022]. However, it remains largely unknown how the effects of
environmental stochasticity and response diversity on population stability and
asynchrony are mediated in the context of complex communities, such as
food-webs.

Complexity-stability theory well documented that the structure of food-webs
determines population stability [@may_will_1972; @angelis_stability_1975;
@mccann_weak_1998; @mccann_reevaluating_1997; @vasseur_environmental_2007;
@eschenbrenner_diversity_2023], and stability-richness relationships
[@thebault_trophic_2005; @thebault_relationship_2006]. Population stability is
indeed documented to decrease with increasing interaction strength
[@may_will_1972; @mccann_weak_1998], connectance [@may_will_1972], but to be
higher at higher trophic level [@shanafelt_stability_2018; @danet_species_2021].
In result, the distribution and strength of species interactions, as well as
trophic level might have a strong effect on community stability and
stability-richness relationships [@thebault_trophic_2005;
@thebault_relationship_2006]. This expectation is reinforced by documented
examples where predators can mediate the effect of response diversity on
asynchrony [@vasseur_environmental_2007], and synchronise the dynamics of their
prey [@raimondo_interspecific_2004; @korpimaki_predatorinduced_2005]. Recent
study in plant communities suggests that asynchrony stems mainly from the
statistical averaging of independent species fluctuations (“portfolio effects”),
rather than compensatory dynamics resulting from response diversity and species
interactions [@zhao_biodiversity_2022]. But we might expect that prey switching
and trophic cascades arising in food-webs lead to higher compensatory dynamics
than in plant communities [@mccann_diversitystability_2000;
@fahimipour_compensation_2017].

Here, we investigated the effects of response diversity and food-web structure
on community stability and on stability-richness relationships in stochastic
environment. To do so, we developed an extension of the Bioenergetic food web
model [@brose_allometric_2006; @lajaaiti_ecologicalnetworksdynamicsjl_2024]
which includes environmental stochasticity and response diversity
[@loreau_species_2008; @vasseur_environmental_2007; @ripa_food_2003]. We
included environmental stochasticity on allometric natural death rate, such as
species of higher trophic level had lower death rate in
average[@vasseur_environmental_2007], response diversity was included as the
inverse of species correlation in their mortality rates.  We first decipher the
respective effects of food-web structure, environmental stochasticity, and
response diversity on community stability and its partition into population
stability and asynchrony (i.e. portfolio and compensatory effects). We then
investigate how food-web structure and response diversity interact to determine
the sign of stability-richness relationship.

```{r}
tar_load(sim_param_tab)
p_mm <- function(
  v = "S",
  data = sim_param_tab
) {
  out <- data[data$Parameter %in% v, ]
  paste0("", signif(out[["min"]], 1), "-", signif(out[["max"]], 1), "")
}
```

```{r, eval = FALSE}
p_mm("ct")
```

```{r}
tar_load(c(sem_doak_eff_tab, sem_simplified_eff_tab))

```{r fig2-coeff, eval = FALSE}
tar_read(sem_fig_coeff)
tar_read(sem_doak_rsquared_csc_rerun)
```

```{r}
tar_load(sim_summary_stat)
m_sim <- function(
  v = "stab_com",
  data = sim_summary_stat,
  col = "Median (5%, 95%)"
  ) {
  data[data$var %in% v, ][[col]]
}
```
```{r, eval = FALSE}
m_sim(v = c("stab_com", "async"))
m_sim(c("stab_com", "async"))
m_sim(col = "n")
```

## Drivers of community stability, population stability and asynchrony

```{r}
get_eff <- function(
  y = "pop_stab", x = "richness",
  d = 2, effect_type = "direct",
  tab = sem_doak_eff_tab) {
  mask <- tab$response == y & tab$predictor == x &
    tab$effect_type == effect_type
  round(tab[mask, ]$effect, d)
}

rsr <- get_eff(y = "sae_doak", x = "richness")
rst <- get_eff(y = "sae_doak", x = "w_avg_tlvl")
rsi <- get_eff(y = "sae_doak", x = "avg_int_strength")
rsc <- get_eff(y = "sae_doak", x = "ct_alive")

rcr <- get_eff(y = "cpe_int", x = "richness")
rci <- get_eff(y = "cpe_int", x = "avg_int_strength")
rct <- get_eff(y = "cpe_int", x = "w_avg_tlvl")
rcd <- get_eff(y = "cpe_int", x = "resp_div")
rcc <- get_eff(y = "cpe_int", x = "ct_alive")

# Detailled SEM coefficients
rsr <- get_eff(y = "sae_doak", x = "richness")
rst <- get_eff(y = "sae_doak", x = "w_avg_tlvl")
rsr <- get_eff(y = "sae_doak", x = "richness")
rsa <- get_eff(y = "sae_doak", x = "avg_int_strength")
rsct <- get_eff(y = "sae_doak", x = "ct_alive")
rsc <- get_eff(y = "sae_doak", x = "c")
rsd <- get_eff(y = "sae_doak", x = "resp_div")

```{r}
rscp <- get_eff(y = "stab_com", x = "pop_stab")
rsca <- get_eff(y = "stab_com", x = "async")

rac <- get_eff(y = "async", x = "cpe")
ras <- get_eff(y = "async", x = "sae_total")
```

The `r format(m_sim(col = "n"), scientific = FALSE)` simulations generated a
wide range of food-web structure (Median (5%, 95%), species richness:
`r m_sim("richness")`, connectance: `r m_sim("ct_alive")`, average interaction
strength: `r m_sim("avg_int_strength")`, maximum trophic level:
`r m_sim("max_tlvl")`). We quantified the  (1) the respective effects of
population
stability and asynchrony on community stability, (2) the source of asynchrony,
either portfolio effects or compensatory dynamics, and disentangled (3) how
stressors (environmental stochasticity, response diversity) and food-web
structure (species richness, average interaction strength, average trophic
level, connectance) determine average population stability, portfolio effects
(using Doak's definition), and compensatory dynamics stemming from species
interactions. We did so by using a structural equation model which included all
the simulation variables as control (i.e. predator interference, predator-prey
mass ratio) to avoid the detection of spurious non-causal effects (See Methods,
Table/Figure SXX). Overall, we found that population stability had a positive
effect twice higher than asynchrony on the temporal stability of community
biomass (resp. $r_\delta$= `r rscp` and `r rsca`, $r_\delta$ being the
standardised coefficient, Fig. \@ref(fig:sem)a). We further found that
asynchrony was driven more by portfolio effects than by compensatory dynamics
stemming from species interactions (resp. $r_\delta$= `r rac` and `r ras`).
However, we found opposite results, i.e. that compensatory dynamics contributed
more than portfolio effects to asynchrony, using an alternative portfolio
effects definition that attributes the effects of response diversity to
compensatory effects instead of portfolio effects (Fig. SXX).

Food-web structure had consistent negative effects on asynchrony, mainly through
portfolio and compensatory effects (Fig. \@ref(fig:sem)a). Indeed, average
trophic level and connectance had negative effects on the compensatory effects
stemming from species interactions (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength had a small negative
effects on portfolio effects ($r_\delta$= `r rsi`). In contrast, response
diversity and species richness had strong positive effects on asynchrony,
mainly through portfolio effects. Response diversity had a strong positive
effect on portfolio effects ($r_\delta$= `r rsd`), and a negative one on
compensatory dynamics ($r_\delta$= `r rcd`). In turn, species richness had a
positive effect on both portfolio and compensatory effects (resp.
$r_\delta$=`r rsr` and `r rcr`). Finally, environmental stochasticity had a
small negative effect on compensatory dynamics.

In turn, we found little effect of food-web structure on population stability,
except for average trophic level having a strong positive effect on
population stability ($r_\delta$= `r rpt`, Fig. \@ref(fig:sem)a). In turn, the
variance of environmental stochasticity had a strong negative effect on
population stability ($r_\delta$= `r rpe`), while response diversity had a small
positive effect on population stability ($r_\delta$= `r rpd`).

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both the variance of environmental stochasticity and species response
diversity had by far the most important effects on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
and average trophic level had similar positive total effects on community stability
($r_\delta$=`r rtsr`), but through opposite pathways and antagonistic effects on
asynchrony and population stability. While species richness increased community
stability through a strong positive effect on asynchrony ($r_\delta$=`r rtar`), it
also weakly decreased population stability ($r_\delta$=`r rtpr`). At the opposite, average
trophic level increased community stability through strong positive effect on
population stability ($r_\delta$=`r rtpt`), but dampened by a negative total
effect on asynchrony ($r_\delta$=`r rtat`). Interestingly, average interaction
strength also had a positive effect on population stability and a negative one
on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`), but resulting in a
total negative effect on community stability ($r_\delta$= `r rtsi`).
Connectance, instead, had a negative effect on community stability ($r_\delta$=
`r rtsct`) through consistent negative effects on asynchrony and population
stability (resp. $r_\delta$=`r rtact` and `r rtpct`).

Using a more complex linear model including interactions among food-web
structure, response diversity and environmental stochasticity, we found that
food-web structure strongly interacts with response diversity to determine the
sign of the stability-richness relationships. In the absence of response
diversity, we found only negative stability-richness, regardless of the food-web
structure (Fig. \@ref(fig:pred)a). Response diversity enabled the rise of
positive stability-richness relationships, enhanced by average interaction
strength and dampened by average trophic level and connectance. Higher
interaction strength led to positive stability-richness relationships for lower
values of response diversity, interaction strength enhanced asynchrony-richness
relationships (Fig. SXX), so that the stronger positive stability-richness
relationships were found at both high interaction strength and response
diversity. In contrast, higher average trophic level and connectance both led to
negative stability-richness relationships, because they led to more negative
population stability-richness relationships and weaker positive
asynchrony-richness relationships (Fig. SXX). Overall, our results highlight
strong contrast between the relatively small effects of food-web structure on overall
community stability (Fig. \@ref(fig:sem)b) and their strong effects on
stability-richness relationships (Fig. \@ref(fig:pred)a).

Temporal community stability is well documented in simplified assemblages
such as plant communities, where community stability is driven by asynchrony in
species fluctuations. In contrast, our results show that community stability in
complex food-webs was mainly driven population stability rather than asynchrony.
In the range of our simulations, we identified that environmental variability
and response diversity were the main driver of community stability, far more
than any food-web structure component, a result consistent across methodological
choices (Fig. SXX). Moreover, we found that species richness and food-web
structure have antagonistic but consistent effect on community stability.
Finally, we found that the sign of the stability-richness relationship is
strongly mediated by the interaction between food-web structure and response
diversity. These results highlight the need to account for both food-web
structure and response diversity to improve our understanding of the stability
of ecological communities in variable environments.

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, which contrasts with empirical
data that reported an equivalent influence of population stability and
asynchrony on community stability in birds and bats assemblages
[@olivier_urbanization_2020]. However, our results are in accordance with
previous findings in empirical food-webs [@danet_species_2021] and
food-web models [@eschenbrenner_diversity_2023]. Are there fundamental
differences between single trophic and food-web in the contribution of
population stability and asynchrony? We found that all food-web metrics, i.e.
average trophic level, connectance, and average interaction strength all
decreased asynchrony. In line with previous studies showing that predators can
synchronise their preys [@raimondo_interspecific_2004;
@kortsch_disentangling_2021], and that predator-prey couples are more
synchronous [@danet_species_2021], those results suggest that trophic links
might dampen overall asynchrony in ecological communities. A complementary
explanation could be related to the biomass structure of food-webs. Food-webs
typically includes range of bodymasses much larger than in species assemblages,
differing in bodymass by several orders of magnitude [@brose_predator_2019].
This highly uneven biomass distribution can translate in uneven species
variability, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits portfolio effects and compensatory
dynamics in species rich communities [@zhao_biodiversity_2022]. The fact that
average trophic level, connectance, and average interaction strength all
decreased evenness of portfolio effects lends toward this hypothesis. Our
results then reinforce the evidence that community stability in food-web might
be driven more by population stability than asynchrony.

We further found that asynchrony was driven more by portfolio effects than
compensatory dynamics. But this results was dependant of the definition of
portfolio effects, as we found the opposite while taking Tilman's
definition[@tilman_ecological_1999; @zhao_biodiversity_2022] instead of Doak's
one[@doak_statistical_1998], in contrast with previous assessment in
experimental plant communities[@zhao_biodiversity_2022]. Portfolio effects is
supposed to measure stabilising effects of species richness in the absence of
species interactions[@loreau_biodiversity_2021]. While Tilman's considers that
portfolio effects are emerging from statistical averaging of independent species
fluctuations alone (i.e. in the absence of covariance among species), Doak's
explicitly states that statistical averaging can only emerge if compensatory
dynamics hold by response diversity is present[@zhao_biodiversity_2022]. While
the definition of portfolio effects is a long-standing
debate[@zhao_biodiversity_2022], our results show that the importance of
portfolio effects depends on where response diversity acts. In agreement with
previous studies[@loreau_species_2008; @doak_statistical_1998;
@thibaut_understanding_2013; @loreau_biodiversity_2021], we believe that the
definition of portfolio effects standing from statistical averaging and
environmental compensatory dynamics generated by response diversity is the more
appropriate as the absence of species interactions does not guarantee independent
species fluctuations, because common environment tend to generate positive
covariances among species[@loreau_biodiversity_2021]. The fact that
environmental stochasticity alone had a really small effect on asynchrony to the
difference of response lends toward this interpretation. So, our study suggests
that response diversity is a key ingredient to generate asynchrony, not only in
plant communities, but also in food-webs. We further show that explicitly
manipulating response diversity highlight that it is the main driver of
asynchrony, and that the respective contribution of portfolio effects and
compensatory dynamics to asynchrony depends on the decision to consider that
response diversity is part of portfolio effects or compensatory dynamics.

We further found that higher average trophic level lead to higher community
stability by increasing average population stability, confirming previous
theoretical[@brose_allometric_2006; @shanafelt_stability_2018;
@eschenbrenner_diversity_2023] and empirical[@danet_species_2021] results. But
this effect disappeared when we simulated community dynamics with equal
mortality rates for all species instead of allometric ones (Fig. S1). It then
suggests that higher stability of food-web with higher average trophic level is
because of species life-history traits and not directly by species interactions.
The assumption that natural mortality rates decreased with bodymass is well
documented across eukaryotes[@hatton_linking_2019] and seems to be a necessary
ingredient for the stabilising effects of higher trophic levels. In turn, we
found that average trophic level and connectance decreases strongly asynchrony
mainly through a negative effect on compensatory dynamics generated by species
interactions. Both higher connectance and average trophic level lead to greater
generalism of predators in the niche model that we used to build food-webs. The
synchronising effects of average trophic level and connectance echo with
empirical findings that a generalist predator can indeed synchronise the
dynamics of numerous preys[@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005]. The overall effect of connectance on overall
community stability was quite low, in agreement with previous empirical
findings[@danet_species_2021]. Similarly, we found very low effects
of the average interaction strength on community stability which in apparent
disagreement with the predominant role of interaction strength and connectance
on stability in the complexity-stability debate [complexity-stability paradox,
@may_will_1972]. But our study focused on the temporal stability of coexisting
species at steady state, which might mask connectance and interaction strength
values that led to species extinctions during the transient phase.

We found that species richness had a positive effect on community stability of
the same importance as average trophic level, but through opposite pathways.
Species richness had strong positive effects on asynchrony through positive
effects on portfolio and compensatory effects. It is in agreement with previous
study showing that species richness can promote asynchrony by both
pathways[@zhao_biodiversity_2022]. Furthermore, our study shows that the positive
effects of species richness on asynchrony is three times lower than the effects
of response diversity. Our model predictions further indicates that in absence
of response diversity, we only observe negative relationships between
community stability and species richness. Therefore, it reinforces the idea that
response diversity is the crucial ingredient that hold the observed
positive effects of species richness on community stability and not species
richness alone[@thebault_trophic_2005; @loreau_species_2008;
@loreau_biodiversity_2021]. In return, species richness had a small negative
effect on average population stability. It is in agreement with previous
meta-analysis reporting that such negative effects are common in natural systems
but weak[@houlahan_negative_2018], but the mechanisms behind this relationship
are still debated. Overall our study highlights that the central role of
response diversity and species richness on asynchrony and thus community
stability also extends to complex food-webs.

Our results further show that food-web structure interact strongly with response
diversity to determine the sign of stability-richness relationships. We found
surprisingly that higher average interaction strength enhance positive
stability-richness relationships, by enhancing asynchrony-richness
relationships. At the opposite, higher average trophic level and higher
connectance both led to negative stability-richness, by dampening
asynchrony-richness relationships. Our result echoes with previous results
showing negative stability-richness relationship in freshwater
food-webs[@danet_species_2021]. Those results also echo with the
complexity-stability debate where population stability decreases as the
community is more species diverse, has higher connectance, and higher
interaction strength[@may_will_1972]. However, we found that the negative
effects of complexity on stability-richness relationships are dampened in
presence of high response diversity. Response diversity might then be one of the
mechanisms that explains why positive relationships between stability and
species richness are so prevalent in empirical settings while food-web
theoretical models that do not often include environmental stochasticity and
response diversity finds negative complexity-stability
relationships[@ives_stability_2007]. It is also in agreement previous studies
showing that higher connectance in low response diversity can drive negative
stability-richness in plant-herbivore systems[@thebault_trophic_2005]. Finally,
found that the strength of environmental stochasticity itself does not
affect the shape of the community stability-richness but rather controls the
overall community stability through its effects on population stability.

In conclusion, our study uncovers complex but consistent effects of response
diversity and food-web structure on the temporal stability of complex
communities. We showed that community stability was more driven by population
stability than asynchrony. We further showed that species richness and average
trophic level both increase community stability but by opposite pathways. Our
study also demonstrates that response diversity is the core mechanism driving
the buffering effects of biodiversity on environmental stochasticity also in
complex food-webs. Finally, our study shows that food-web structure and response
diversity interact to drive a continuum from negative to positive
stability-richness relationships. Looking to practical applications, our study
highlights that species richness without response diversity does not lead to
positive community-richness relations. So that assessment on how ecological
communities widespread recomposition[@dornelas_assemblage_2014;
@blowes_geography_2019] and homogenisation[@olden_homogocene_2018;
@muthukrishnan_invasive_2020; @wang_biotic_2021] is affecting response diversity
would be crucial to accurately predict long-term consequences on community
stability. The second take away to be drawn is that the variance of
environmental stochasticity was the main driver of community stability,
suggesting that increase in species richness and the maintenance of high
response diversity would be not enough to compensate for strong increase in
environmental variability in the coming decades[@lee_ipcc_2023].

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017]. This model
is widely used to simulate the dynamics of complex ecological communities
because of the simplicity  and because it can be parametrized using  the
metabolic theory of ecology [@brown_toward_2004]. The model describes species
biomass dynamics over time with two master equations, one for the primary
producers (eq. \@ref(eq:prod)) and one for the consumers (eq.\@ref(eq:cons)):

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i - d_i B_i
\end{equation}

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
    - d_i B_i
\end{equation}

@@

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + c_i B_i
    	+ \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

@@

@@
